---
title: "balafoutas_exercise"
author: "Daniel Toribio-Flórez"
date: "15/02/2021"
output: html_document
---

Load packages:
```{r}
library(tidyverse)
library(paramtest)
library(parallel)
```

# Balafoutas et al. (2014):
Observed frequencies of Third-Party Punishment (3PP) in Experimental Conditions:

```{r}
data <- matrix(c(20,20,36,4), nrow = 2, 
               dimnames = list(c("No Punish","Punish"),
                       c("No Counterpunishment","Counterpunishment")))
```

Balafoutas et al. performed a Fisher's exact test to test whether the frequency of 3PP differed across conditions:

```{r}
fisher.test(data)
```

**Main Result:** The decision to exert 3PP significantly depend on whether or not there is a risk of being counterpunished.

# Power Simulation:
We base our power simulation on the reported effect size (i.e., difference in frequencies).

## Parameters based on Balafoutas et al. (2014):

Sample Size: N = 80 (40 per cell)
Probability of 3PP in "No Counterpunishment" condition: prob1 = 20/40 = .5
Probability of 3PP in "Counterpunishment" condition: prob2 = 4/40 = .1

## Simulation function:

```{r}
fisher_sim <- function(simNum,N,prob1,prob2){ #necessary parameters.
  
cond <- rep(0:1, N/2) #Experimental Condition. 0 = No Counterpunishment, 1 = Counterpunishment.
df <- as.data.frame(cond) #Create data set.
df$punish <- ifelse(cond == 0, rbinom(N/2,1,prob = prob1),
                    rbinom(N/2,1,prob = prob2)) #Conditional binomial distribution based on Condition.
tab <- table(df$cond, df$punish) #Create contingency table.

test <- fisher.test(tab) #Run Fisher's exact test 
#Consider to change to Chi-squared with bigger samples)
# test <- chisq.test(tab)

p <- test$p.value #Extract p-value from test output.
sig <- p < .05 #Returns TRUE if p-value < alpha.
return(c(p = p, sig = sig))
}
```

## Parallel Iterations:

Detect number of cores that your computer can use for performing parallel iterations.
```{r}
detectCores()
```

Perform power simulation with 1000 iterations:
```{r}
set.seed(1234) #Set seed to reproduce same results.
power_log <- grid_search(fisher_sim,
                         params = list(N = 80,
                                       prob1 = .5,
                                       prob2= .1),
                         n.iter = 1000,
                         output = "data.frame",
                         parallel = "snow",
                         ncpus = 8)
```

See results:
```{r}
power_log
```

Calculate Statistical Power:
```{r}
results(power_log) %>%
  group_by(N.test) %>%
  summarise(
    power = mean(sig)
  )
```

# EXERCISES

We are interested in adding a condition to Ballafoutas experimental setup.
We expect that in this new condition we will observe a lower frequency of punishment
than in "0 - no counterpunishment", but greater than in "1 - counterpunishment".

Thus, one first step is to check whether the sample size used by Balafoutas et al. (N = 80)
would allow us to capture smaller differences between conditions with sufficient statistical power.

## 1. ##
Use the "fisher_sim" function we created to estimate the smallest difference we can capture
with 90% power, given N = 80.

```{r}

```

```{r}

```

In case you wanted to plot your results:
```{r}

```


## 2. ##
Consider that we want to be able to capture differences of minimum .20,
with at least 90% power. How much larger should our sample size be?

CLUE 1: In this case, it will be more than a few, so think big!
CLUE 2: Since sample sizes will get big, it would be recommendable to first
change the Fisher's exact test for Chi-squared test in the simulation function, 
because the latter performs better with big sample sizes. You can use the `chisq.test()`
function for this purpose.

### Simulation function:

```{r}

```

### Parallel Iterations:

```{r}

```

### Results:
```{r}

```

###Calculate Statistical Power:
```{r}

```

In case you want to plot your results:
```{r}

```


