---
title: "first tutorial"
author: "Niklas Cypris"
date: "19 2 2021"
output: pdf_document
---

```{r setup}
library(tidyverse)
library(paramtest)
library(parallel)
```

# SIMULATION FUNCTION

y = b0 + b1 * x + error

```{r}
set.seed(161)
lm_sim <- function(simNum, N, b1, b0 = 0) #necessary parameters
{
  
  con <- # create normally distributed data using rnorm() function
  rerror <- # squrt(variance of pro-environmental behavior that is not predicted by conscientiousness)
  y <- # formula: intercept + slope*conscientiousness + random error
  model1 <- lm() #predict pro-environmental behavior from conscientiousness lm()
  
  # model 1
  est1_x1 <- coef(summary(model1))['con', 'Estimate']
  p1_x1 <- coef(summary(model1))['con', 'Pr(>|t|)']
  sig1_x1 <- p1_x1 < .05
  
  return(c(est1_x1 = est1_x1, p1_x1 = p1_x1, sig1_x1 = sig1_x1))
  
}
```

## PARALLEL ITERATIONS

Detect number of cores that your computer can use for performing parallel iterations.

```{r}
detectCores()
```

# SIMULATE DATA

```{r}
power_lm <- grid_search(lm_sim, params=list(N = c(150, 200, 250, 300)),
                         n.iter = 1000, output ='data.frame', b1 = .2, parallel = 'snow', ncpus = 4)
```

# DISPLAY RESULTS

```{r}
power_lm %>%
  results() %>% 
  group_by(N.test) %>%
  summarise(
    beta_con = mean(est1_x1),
    power = mean(sig1_x1)
  )
```

```{r}
power_lm %>% 
  results() %>% 
  group_by(N.test) %>% 
  summarise(
    beta_con = mean(est1_x1),
    power = mean(sig1_x1)
  ) %>% 
  ggplot(aes(x = N.test, y = power)) +
  geom_line() +
  theme_classic()
```

```{r}
power_lm %>% 
  results() %>% 
  ggplot(aes(x = round(p1_x1, 2))) +
  geom_bar() +
  geom_vline(xintercept = .05, color = "red") +
  facet_wrap(~ N.test, nrow = 1) +
  xlab("p values") +
  xlim(-0.01, 0.5) +
  theme_classic()
```

# EXERCISES

1. Vary the effect size of the b1-coefficient (.1, .2, .3) and the number of participants (50, 100, 150).

```{r}

```

2. Add another predictor to the equation. Adjust output (e.g. est1_x2 etc) and results table

```{r}

```

3. Compare model1 with one predictor (b1 = .05) and model2 with two predictors (b1 = .05, b2 = .2, bx = .05).

```{r}

```

4. Simulate Data for a t-test.

```{r}

```

